ASYMPTOTIC VALIDITY OF THE BAYES-INSPIRED INDIFFERENCE ZONE PROCEDURE:

### 2015-05-07 paper/wish/wsc15paper.tex
meeting
changes to text during meeting

\section*{ABSTRACT}
We consider the indifference-zone (IZ) formulation of the ranking and selection problem in which the goal is to choose an alternative with the largest mean with guaranteed probability, as long as the difference between this mean and the second largest exceeds a threshold.
Conservatism leads classical IZ procedures to take too many samples in problems with many alternatives. The Bayes-inspired
Indifference Zone (BIZ) procedure, proposed in \citeN{Frazier:BIZ}, is
less conservative than previous procedures, but its proof of validity
requires strong assumptions, specifically that samples are normal, and variances are known 
with an integer multiple structure. In this paper, 
we show asymptotic validity of a slight modification of the original BIZ procedure 
as the difference between the best alternative and the second best goes to zero,
when the variances are known and finite, and samples are independent and identically distributed, but not necessarily normal.

@@

\citeN{Frazier:BIZ} eliminates the use of Bonferroni's inequality, reducing conservatism. This paper presents a new sequential elimination
IZ procedure, called BIZ (Bayes-inspired Indifference Zone), whose
lower bound on worst-case probability of correct selection in the
preference zone is tight in continuous time, and almost tight in the
discrete time. In numerical experiments, the number of samples required
by BIZ is significantly smaller than that of procedures like the KN procedure of \citeN{Kim:2001} and
the $\mathcal{P}_{B}^{*}$ procedure of \citeN{Bechhofer:1968}, 
especially on problems with many alternatives. Unfortunately, the
proof that the BIZ procedure satisfies the IZ guarantee for the discrete-time
case assumes that variances are known and have an integer multiple
structure which is not very realistic.

The contribution of this work is that we prove the asymptotic validity of the 
BIZ procedure for the discrete-time case when the variances are known and $\delta$ goes to zero. 
In order to simplify analysis, we consider a slightly modification of this procedure.
Although we only proved it in the known variance setting, it is still a significantly
generalization over the assumption that the variances are known and have an integer multiple
structure. Furthermore, we do not need to assume that the alternatives follow a normal distribution. 
The only assumptios are that the alternatives are independent, identically distributed and have
finite variance. We conjecture that these techniques can be used to show the validity of this procedure when
the variances are unknown, and we present numerical experiments that support this belief. 
\citeN{kim:asympt} also proves the asymptotical validity of a IZ procedure in the same limit.
Our proof shares some similarities since we both use a central limit theorem.

Two other common sources of conservatism are the change from discrete-time processes to continuous-time processes used to show IZ guarantees, and the fact that typically, the configuration under consideration is not a worst-case configuration \cite{Kim:Conserv}.  
This last source tends to be the largest in terms of its contribution to conservatism, and the first tends to be the smallest, with the contribution from Bonferonni inequality's lying between \cite{Kim:Conserv}.
Although the BIZ procedure avoids the Bonferonni inequality, it does not avoid these other two sources of conservatism.
Indeed, all indifference zone procedures must meet the PCS guarantee for all configurations in the preference zone, including worst-case configurations.  In this sense, this last source of conservatism arises from the indifference zone guarantee itself, rather than the from a design of a particular procedure.

@@

\section{THE BAYES-INSPIRED IZ (BIZ) PROCEDURE}

@@

% Using that $\mathcal{C}\left(\delta,\cdot\right)$ is right-continuous,
% we can prove that the difference of the probability of correct selection of the discrete-time BIZ algorithm and the  probability of correct selection of the continuous-time BIZ algorithm converge to zero as $\delta$ goes to zero. Then we can use the continuous-time BIZ algorithm in the limit, i.e. we can use $\mathcal{C}\left(\delta,\tau_{n}\left(\delta\right)\right)$ instead
% of $\mathcal{C}\left(\delta,\hat{\tau}_{n}\left(\delta\right)\right)$. The details may be found in the full version of this paper \cite{fullpaper}. 
The following lemma allows us to use the continuous-time BIZ algorithm in the limit. A proof Lemma~\ref{l:interpolation} may be found in a full version of this paper \cite{fullpaper}.

\begin{lemma}
If samples from system $x\in\left\{ 1\ldots,k\right\} $ are identically
distributed with finite variance and independent, over time and across alternatives, then
\begin{equation*}
    \lim_{\delta\rightarrow0} \mathbb{P}\left( h\left( \left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right) \right) =k\right) = \lim_{\delta\rightarrow0}
 \mathbb{P}\left(
 f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1
 \right).
\end{equation*}

@@

\begin{theorem}
If samples from system $x\in\left\{ 1\ldots,k\right\} $ are identically
distributed and independent, over time and across alternatives, then
$\mbox{lim}_{\delta\rightarrow0}\mbox{PCS}(\delta) \geq P*$
provided $\mu_{k}=a_{k}\delta,\mu_{k-1}=a_{k-1}\delta,\ldots,\mu_{1}=a_{1}\delta$,
$a_{k}>a_{k-1}\geq \cdots \geq a_{1}$, $a_{k}-a_{k-1}\ge1$, and the variances are finite and
do not depend on $\delta$.

@@

\begin{proof}
Using the definitions given at the beginning of this section,
the selection decision of the discrete-time BIZ procedure for a particular $\delta>0$ when $\mu_{k}=a_{k}\delta,\mu_{k-1}=a_{k-1}\delta,\ldots,\mu_{1}=a_{1}\delta$ is given by
\begin{equation*}
h\left(\left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right)\right)
\end{equation*}
and the probability of correct selection $\mbox{PCS}(\delta)$ is
\begin{equation*}
    \mbox{PCS}(\delta) =  \mathbb{P}\left( h\left( \left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right) \right) =k\right).
\end{equation*}
By Lemma~\ref{l:interpolation}, as $\delta\to0$, we have that 
\begin{equation}
\lim_{\delta\to0}
\mbox{PCS}(\delta)
% P\left( h\left( \left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right) \right) =k\right)
 = \lim_{\delta\rightarrow0}
 \mathbb{P}\left(
 f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1
 \right)
 \label{eq:convergence1}
\end{equation}

We also have, by Lemma~\ref{l:continuity} and an extension of the continuous mapping theorem (Theorem~5.5 of \citeN{billingsley:convergence}), 
\begin{equation*}
    f\left(\mathcal{C}\left(\delta,t\right),\delta\right)\Rightarrow g\left(W\left(t\right)\right)
%\label{eq:convergence}
\end{equation*}
in distribution as $\delta\rightarrow0$.
This implies that
\begin{equation}
\lim_{\delta\rightarrow0}\mathbb{P}\left(f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1\right)
 = \mathbb{P}\left(g\left(W\right)=1\right)
 \label{eq:convergence2}
\end{equation}

@@

We show the performance of the BIZ procedure in discrete time with maximum elimination
($c=1-(P^*)^\frac{1}{k-1}$) when $\delta$ goes to zero. We plot the PCS of three different examples.

First, we consider the known variance case. Specifically, we consider a slippage configuration 
with $100$ systems normally distributed, where $\mu_{k}=\delta,\mu_{k-1}=0,\dots,\mu_{1}=0$, $\delta$ is 
within the interval $[0.1,10]$, and $\lambda_{100}=1,\lambda_{99}=1+\frac{(0.5)(98)}{99},\cdots,\lambda_{1}=0.5$. Here, $P^*=0.9$ and $n_{0}=0$.
Figure ~\ref{fig: tahi3} shows that in this case the IZ guarantee is always satisfied, and the bound on probability of correct selection is tight when $\delta$ goes to zero.
When $\delta$ is big enough, the PCS is almost one because the difference between the best system and the others is large enough to be easily identifiable by the
BIZ procedure. Here, the points plotted have central confidence intervals of length at most $0.014$.

\begin{figure}[h]
  \centering
   \begin{subfigure}[b]{0.3\textwidth}
      \centering
       \includegraphics[width=0.7\textwidth]{plot5}
    \caption{
    \label{fig: tahi3}}
   \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
         \includegraphics[width=0.7\textwidth]{plot3}
             \caption{
    \label{fig: tahi2}}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
         \includegraphics[width=0.7\textwidth]{plot1}
    \caption{ \label{fig: tahi}}
    \end{subfigure}
    \caption{(a)  Known variances and $P^*=0.9$. In this example, our theorem is true: the IZ guarantee is always satisfied and the inequality is tight as $\delta$ goes to zero. (b) Unknown variances and $P^*=0.9$. In this example, our theorem is true: the IZ guarantee is satisfied and the inequality is tight as delta goes to zero. If we want that the IZ guarantee be satisfied in all the cases, we can just increment $n_{0}$. (c) Known variances, $P^*=0.9$ This was a hard example to find and it was specially chosen to illustrate our theorem. In practice, we would choose $n_{0}>0$ and the IZ guarantee will be satisfied.
\label{fig:tahi10}}

\end{figure}

We now consider the unknown variance case. Specifically, we consider a slippage configuration 
with $100$ systems normally distributed, where $\mu_{100}=\delta,\mu_{99}=0,\dots,\mu_{1}=0$, $\delta$ is 
within the interval $[0.1,10]$, and $\lambda_{100}=10,\lambda_{99}=\cdots=\lambda_{1}=1$. Here, $P^*=0.9$ and $n_{0}=15$.
Figure ~\ref{fig: tahi2} gives us evidence that our theorem should also be true when the variances are unknown. We should note that
this is a difficult example, and still we have a good performance in most of the cases. However, if we want to always satisfy the IZ guarantee, 
we can just increment the parameter $n_{0}$. Here, the points plotted have central confidence intervals of length at most $0.014$.

@@

\noindent {\bf PETER I. FRAZIER} is an Assistant Professor of the School of Operations Research and Information Engineering at Cornell University, Ithaca, NY.  He holds a Ph.D. in operations research and financial engineering industrial engineering from Princeton University. His research interests include optimal learning, sequential decision-making under uncertainty, and machine learning, focusing on applications in simulation optimization, design of experiments, materials science, e-commerce and medicine. He is a member of INFORMS, and SIAM. His e-mail address is \email{pf98@cornell.edu}.\\

>>>>

\section*{ABSTRACT}
We consider the indifference-zone (IZ) formulation of the ranking and selection problem in which the goal is to choose an alternative with the largest mean with guaranteed probability, as long as the difference between this mean and the second largest exceeds a threshold.
Conservatism leads classical IZ procedures to take too many samples in problems with many alternatives. The Bayes-inspired
Indifference Zone (BIZ) procedure, proposed in Frazier (2014), is
less conservative than previous procedures, but its proof of validity
requires strong assumptions, specifically that samples are normal, and variances are known 
with an integer multiple structure. In this paper, 
we show asymptotic validity of a slight modification of the original BIZ procedure 
as the difference between the best alternative and the second best goes to zero,
when the variances are known and finite, and samples are independent and identically distributed, but not necessarily normal.

@@

In addition to Bonferroni's inequality, two other common sources of conservatism in indifference-zone ranking and selection procedures are the change from discrete time to continuous time often used to show IZ guarantees, and the fact that typically, the configuration under consideration is not a worst-case configuration \cite{Kim:Conserv}.  The difference between worst and typical cases tends to contribute the most to conservatism, with Bonferonni's inequality contributing second-most, and the continuous/discrete time difference contributing the least \cite{Kim:Conserv}.
Although the difference between the worst and typical cases is the largest contributor to conservatism, all indifference zone procedures must meet the PCS guarantee for all configurations in the preference zone, including worst-case configurations, and so this source of conservatism is fundamental to the indifference-zone formulation.  Thus, eliminating the use of Bonferroni's inequality remains an important route for reducing conservatism while still retaining the indifference-zone guarantee.

\citeN{Frazier:BIZ} presents a new sequential elimination
IZ procedure, called BIZ (Bayes-inspired Indifference Zone), 
that eliminates the use of Bonferroni's inequality, reducing conservatism. 
This procedure's lower bound on worst-case probability of correct selection in the
preference zone is tight in continuous time, and almost tight in the
discrete time. In numerical experiments, the number of samples required
by BIZ is significantly smaller than that of procedures like the KN procedure of \citeN{Kim:2001} and
the $\mathcal{P}_{B}^{*}$ procedure of \citeN{Bechhofer:1968}, 
especially on problems with many alternatives. Unfortunately, the
proof from \cite{Frazier:BIZ} that the BIZ procedure satisfies the IZ guarantee for the discrete-time case assumes that (1) samples are normally distributed; 
(2) variances are known;
and (3) the variances are either common across alternatives, or have an unrealistic integer multiple structure.

The contribution of this work is to prove the asymptotic validity of the BIZ procedure as $\delta$ goes to zero, retaining the assumption of known variances, but replacing assumptions (1) and (3) by the much weaker assumption of independent and identically distributed finite variance samples.  Thus, our proof allows a much broader set of sampling distributions than that allowed by \cite{Frazier:BIZ}, including non-normal samples and general heterogeneous variances. 
We also show that this bound on worst-case PCS is asymptotically tight as $\delta$ goes to zero, 
showing that the BIZ procedure successfully eliminates conservatism due to Bonferonni's inequality in this more general setting, just as was demonstrated by \citeN{Frazier:BIZ} for more restricted settings.

To simplify our analysis, we analyze a slight modification of the version of the BIZ procedure presented in \cite{Frazier:BIZ}, which keeps a certain parameter $\lambda^2_z$ fixed rather than letting it vary as did \cite{Frazier:BIZ}.  Numerical experiments on typical cases show little difference in performance between the version of BIZ we analyze and the version in \cite{Frazier:BIZ}.
We conjecture that a proof technique similar to the one presented here can be used to show asymptotic validity of the BIZ procedure when
the variances are unknown, and we present numerical experiments that support this belief. 
% \citeN{kim:asympt} also proves the asymptotical validity of a IZ procedure in the same limit.
% Our proof shares some similarities since we both use a central limit theorem.

@@

\section{THE BAYES-INSPIRED IZ (BIZ) PROCEDURE}
\label{sec:BIZ}

@@

% Using that $\mathcal{C}\left(\delta,\cdot\right)$ is right-continuous,
% we can prove that the difference of the probability of correct selection of the discrete-time BIZ algorithm and the  probability of correct selection of the continuous-time BIZ algorithm converge to zero as $\delta$ goes to zero. Then we can use the continuous-time BIZ algorithm in the limit, i.e. we can use $\mathcal{C}\left(\delta,\tau_{n}\left(\delta\right)\right)$ instead
% of $\mathcal{C}\left(\delta,\hat{\tau}_{n}\left(\delta\right)\right)$. The details may be found in the full version of this paper \cite{fullpaper}. 
The following lemma shows that the difference in the correct selection events obtained from applying the BIZ selection map $h$ to the discrete-time and continuous-time versions of $q_{tx}(A)$ vanish as $\delta$ goes to $0$.
A proof of Lemma~\ref{l:interpolation} may be found in a full version of this paper \cite{fullpaper}.

\begin{lemma}
% If samples from system $x\in\left\{ 1\ldots,k\right\} $ are identically
% distributed with finite variance and independent, over time and across alternatives, then
    $\lim_{\delta\rightarrow0} \mathbb{P}\left( h\left( \left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right) \right) =k\right) = \lim_{\delta\rightarrow0}
 \mathbb{P}\left(
 f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1
 \right)$.

@@

\begin{theorem}
\label{t:main}
If samples from system $x\in\left\{ 1\ldots,k\right\} $ are identically
distributed and independent, over time and across alternatives, then
$\mbox{lim}_{\delta\rightarrow0}\mbox{PCS}(\delta) \geq P*$
provided $\mu_{k}=a_{k}\delta,\mu_{k-1}=a_{k-1}\delta,\ldots,\mu_{1}=a_{1}\delta$,
$a_{k}>a_{k-1}\geq \cdots \geq a_{1}$, $a_{k}-a_{k-1}\ge1$, and the variances are finite and
do not depend on $\delta$.

@@

\begin{proof}
Using the definitions given at the beginning of this section,
the selection decision of the discrete-time BIZ procedure for a particular $\delta>0$ when $\mu_{k}=a_{k}\delta,\mu_{k-1}=a_{k-1}\delta,\ldots,\mu_{1}=a_{1}\delta$ is given by
\begin{equation*}
h\left(\left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right)\right)
\end{equation*}
and the probability of correct selection $\mbox{PCS}(\delta)$ is
\begin{equation*}
    \mbox{PCS}(\delta) =  \mathbb{P}\left( h\left( \left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right) \right) =k\right).
\end{equation*}
By Lemma~\ref{l:interpolation}, we have that 
\begin{equation}
\lim_{\delta\to0}
\mbox{PCS}(\delta)
% P\left( h\left( \left(q_{tx}^{'}\left(A\right):t\in\mathbb{N},A\subset\left\{ 1,\ldots,k\right\} ,x\in A\right) \right) =k\right)
 = \lim_{\delta\rightarrow0}
 \mathbb{P}\left(
 f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1
 \right).
 \label{eq:convergence1}
\end{equation}

We also have, by Lemma~\ref{l:continuity} and an extension of the continuous mapping theorem (Theorem~5.5 of \citeN{billingsley:convergence}), 
\begin{equation*}
    f\left(\mathcal{C}\left(\delta,t\right),\delta\right)\Rightarrow g\left(W\left(t\right)\right)
%\label{eq:convergence}
\end{equation*}
in distribution as $\delta\rightarrow0$.
This implies that
\begin{equation}
\lim_{\delta\rightarrow0}\mathbb{P}\left(f\left(\mathcal{C}\left(\delta,t\right),\delta\right)=1\right)
 = \mathbb{P}\left(g\left(W\right)=1\right).
 \label{eq:convergence2}
\end{equation}

@@

We now use simulation experiments to illustrate and further investigate the phenomenon characterized by Theorem~\ref{t:main}. Using the version of BIZ described in Section~\ref{sec:BIZ} with maximum elimination ($c=1-(P^*)^\frac{1}{k-1}$), we estimate and then plot the PCS as a function of $\delta$ using 10,000 independent replications. We consider three different examples, setting $P^*=0.9$ in all three.

Our first example, illustrated in Figure~\ref{fig: tahi3}, is a known variance slippage configuration. Specifically, we consider $100$ systems with independent normally distributed samples, where $\mu_{k}=\delta,\mu_{k-1}=0,\dots,\mu_{1}=0$, $\delta$ is within the interval $[0.1,10]$, and $\lambda_{100}=1,\lambda_{99}=1+\frac{(0.5)(98)}{99},\cdots,\lambda_{1}=0.5$. Here, $n_{0}=0$. Figure ~\ref{fig: tahi3} shows that in this example the IZ guarantee is always satisfied, and the PCS approaches $P^*$ as $\delta$ goes to zero. When $\delta$ is big enough, the PCS is almost one because the difference between the best system and the others is large enough to be easily identifiable by the BIZ procedure. Here, the points plotted have central confidence intervals of length at most $0.014$.

\begin{figure}[h]
  \centering
   \begin{subfigure}[b]{0.3\textwidth}
      \centering
       \includegraphics[width=\textwidth]{plot5}
    \caption{
    \label{fig: tahi3}}
   \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
         \includegraphics[width=\textwidth]{plot3}
             \caption{
    \label{fig: tahi2}}
    \end{subfigure}
        \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
         \includegraphics[width=\textwidth]{plot1}
    \caption{ \label{fig: tahi}}
    \end{subfigure}
    \caption{(a)  Known variances and $P^*=0.9$. In this example, our theorem is true: the IZ guarantee is always satisfied and the inequality is tight as $\delta$ goes to zero. (b) Unknown variances and $P^*=0.9$. In this example, our theorem is true: the IZ guarantee is satisfied and the inequality is tight as delta goes to zero. If we want that the IZ guarantee be satisfied in all the cases, we can just increment $n_{0}$. (c) Known variances, $P^*=0.9$ This was a hard example to find and it was specially chosen to illustrate our theorem. In practice, we would choose $n_{0}>0$ and the IZ guarantee will be satisfied.
\label{fig:tahi10}}
\end{figure}

Our second example, illustrated in Figure~\ref{fig: tahi2}, is an unknown variance slippage configuration. Although Theorem~\ref{t:main} applies only to the known-variance version of BIZ, we conjecture that the unknown-variance version of BIZ should exhibit similar behavior.
In this example, we consider $100$ systems with independent normally distributed samples, where $\mu_{100}=\delta,\mu_{99}=0,\dots,\mu_{1}=0$, $\delta$ is within the interval $[0.1,10]$, and $\lambda_{100}=10,\lambda_{99}=\cdots=\lambda_{1}=1$. We set $n_{0}=15$.
In this example, we have intentionally chosen $n_0$ to be small, and have chosen a large variance for the best system, to cause BIZ to fail to meet the IZ guarantee for $\delta>0$.

We should note that
this is a difficult example, and still we have a good performance in most of the cases. However, if we want to always satisfy the IZ guarantee, 
we can just increment the parameter $n_{0}$. Here, the points plotted have central confidence intervals of length at most $0.014$.

@@

\noindent {\bf PETER I. FRAZIER} is an Assistant Professor of the School of Operations Research and Information Engineering at Cornell University, Ithaca, NY.  He holds a Ph.D. in operations research and financial engineering industrial engineering from Princeton University. His research interests include optimal learning, sequential decision-making under uncertainty, and machine learning, focusing on applications in simulation optimization, design of experiments, materials science, e-commerce and medicine. He is the secretary of the INFORMS Simulation Society.
His e-mail address is \email{pf98@cornell.edu}.\\
